# Reproducible Research: Peer Assessment 1

<<<<<<< HEAD
## Loading and preprocessing the data
```{r}
PA1RR <- read.csv("~/Classes/MOOC/Johns Hopkins Course Materials/Reproducible Research/Data/activity.csv")
str(PA1RR)
```

## What is mean total number of steps taken per day?

First aggregate the data
```{r}
PA1stepsum <- aggregate(steps~date,PA1RR,sum)
head(PA1stepsum)
```
`1.` Histogram of the total number of steps taken in a day

```{r}
hist(PA1stepsum$steps,xlab="Total Number of Steps Taken in a Day",
     main="Histogram of the Total Number of Steps Taken in a Day",
     col="light blue")
```

`2.` The mean umber of steps taken per day is `r mean(PA1stepsum$steps)` and the median is `r median(PA1stepsum$steps)`.  The code used for these in line commands was  

```
mean(PA1stepsum$steps)  
median(PA1stepsum$steps)
```

## What is the average daily activity pattern?

Get the data in the proper form

```{r}
PA1stepave <- aggregate(steps~interval,PA1RR,mean)
head(PA1stepave)
```

`1.`  Time series plot

```{r}
plot(steps~interval,PA1stepave,type="l",xlab="Interval",ylab="Average Number of Steps",
     main="Average Number of Steps Per Day for Each Interval")
```

`2.` The interval with the largest average number of steps is `r PA1stepave$interval[which.max(PA1stepave$steps)]`.  The code used for this is  
```
PA1stepave$interval[which.max(PA1stepave$steps)]
```
## Imputing missing values

`1.` The total number of missing values in the data set is 2304.

```{r message=FALSE}
library(Hmisc)
describe(PA1RR)
```


`2.` Will impute the values using the mean for the corresponding interval across all days.  

`3.` We will augment our data frame with a column indicated the imputed value and then will be have another column with the potential imputed value.  
```{r}
# Create an Imputed column
PA1RRimpute=cbind(PA1RR,Imputed=is.na(PA1RR$steps))
# Merge to get the interval averaged values
PA1RRimpute=merge(PA1RRimpute,PA1stepave,by=c("interval"))
# Place imputed values in
PA1RRimpute$steps.x[PA1RRimpute$Imputed]<-PA1RRimpute$steps.y[PA1RRimpute$Imputed]
# Sort the data back into its original form
PA1RRimpute <- PA1RRimpute[order(PA1RRimpute$date,PA1RRimpute$interval),]
# Clean up
names(PA1RRimpute)[2]="steps"
```


`4.` Histogram
```{r}
PA1stepsumimpute <- aggregate(steps~date,PA1RRimpute,sum)
```  

```{r}
hist(PA1stepsumimpute$steps,xlab="Total Number of Steps Taken in a Day (Imputed Values)",
     main="Histogram of the Total Number of Steps Taken in a Day",
     col="light blue")
```  

The mean and median total number of steps taken per day. day is `r mean(PA1stepsumimpute$steps)` and the median is `r median(PA1stepsumimpute$steps)`.  The code used for these in line commands was  

```
mean(PA1stepsumimpute$steps)  
median(PA1stepsumimpute$steps)
```

The imputing made little change in the mean and median, in fact the mean is unchanged and the median is only slightly larger.  

## Are there differences in activity patterns between weekdays and weekends?

First, we must get the data in a form we can use.  We will create a new column with the weekday in it.  

```{r}
PA1RRimpute <- cbind(PA1RRimpute,Weekday=weekdays(as.Date(PA1RRimpute$date)))
```

`1.` Now create the Weekday/Weekend variable

```{r}
PA1RRimpute <- cbind(PA1RRimpute,
        Work=ifelse(PA1RRimpute$Weekday%in%c("Saturday","Sunday"),"Weekend","Weekday"))
```

`2.` Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).

Load lattice and then aggregate data.

```{r message=FALSE}
library(lattice)
PA1stepaveWork <- aggregate(steps~interval:Work,PA1RRimpute,mean)
head(PA1stepaveWork)
```

```{r}
xyplot(steps~interval|Work,PA1stepaveWork,type="l",
       xlab="Interval",ylab="Average Number of Steps",
     main="Average Number of Steps Per Day for Each Interval by Work Week",
     layout=c(1,2))
```
